---
title: "Clustering"
author: "Connor"
date: "5/21/2023"
output: html_document
---


Goal of this Rmd: Cluster countries based on ND-GAIN Data

Steps:
- Reshape data
- Impute missing Data
- Determine optimal # of clusters
- Cluster data
- Analyze results

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)
library(dplyr)
library(missRanger)
library(cluster)
library(factoextra)
library(randomForest)
library(ClustImpute)
library(plotly)
#library(estlibrary)

options(scipen = 99)
options(ggrepel.max.overlaps = Inf)
set.seed(3)
```

```{r}
nd_gain_data <- read_csv(here("data/nd-gain-data/joined_data/nd-gain-joined.csv"))
```

```{r}
nd_gain_data$value <- round(nd_gain_data$value, digits = 2)
```

```{r}
nd_gain_data <- nd_gain_data %>% 
  filter(year == 2020)
```

```{r}
nd_gain_data <- drop_na(nd_gain_data)
```

```{r}
nd_gain_data <- nd_gain_data %>% 
  pivot_wider(names_from = indicator, 
              values_from = value)
  
```


```{r}
nd_gain_data <- nd_gain_data %>% 
  select(-ISO3, -year, -sids_context, -population)
```


```{r}
nd_gain_data <- nd_gain_data %>% 
  remove_rownames %>% 
  column_to_rownames(var="Name")
```


```{r}
nd_gain_data_less_na <- nd_gain_data[, which(colMeans(!is.na(nd_gain_data)) > 0.6)]



#sum(is.na(nd_gain_data_less_na))/prod(dim(nd_gain_data_less_na)) # percentage of data we will be imputing based on number entered above
```



Impute using ClustImpute

* For nr_cluster = x, we are determining the number of clusters that we believe there are in the data
  - This number can be changed and has significant influence on the output data

```{r}
res <- ClustImpute(nd_gain_data_less_na, nr_cluster = 4)
```


Go into the ClustImpute object and retrieve the imputed data frame

```{r}

nd_gain_data_Clust_Impute <- res$complete_data

```


This visualization helps us to determine the number of clusters that should be used in the kmeans() function

```{r}
fviz_nbclust(nd_gain_data_Clust_Impute, kmeans, method = "silhouette")
```

Remove columns with 0 variance because columns with 0 variance cause an error in clustering

```{r}
nd_gain_data_Clust_Impute <- nd_gain_data_Clust_Impute[, sapply(nd_gain_data_Clust_Impute, var) != 0]
```


Cluster the newly imputed data
```{r}
set.seed(3)
kmeans2 <- kmeans(nd_gain_data_Clust_Impute, center = 3, nstart = 100)
```

Visualize clusters
```{r}
fviz_cluster(kmeans2, data = nd_gain_data_Clust_Impute,
geom = c("text","point"),
             repel = TRUE,            # Avoid label overlapping
             show.clust.cent = TRUE, # Show cluster centers
             palette = "lancet",         # Color palette see ?ggpubr::ggpar
             ggtheme = theme_bw()+
              theme(axis.text.x = element_text( hjust = 0.5, vjust = 0.5, size=15),
        axis.title.x =element_text(size=20),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5, size=15),
        axis.title.y =element_text(size=20),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"),
        strip.text = element_text(size=25)))
```

```{r}
nd_gain_data_Clust_Impute_clusters <- as.data.frame(kmeans2$cluster)
```


